---
title: "PCA plot of tumor-normal paired samples using the vst expression of all available genes"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Reads in total RNA reads count data
```{r,message=FALSE,warning=FALSE}
library(DESeq2)
rna <- read.csv("phase1234_totalRNA_reads_count_ready.csv",header=T,row.names=1)
```

# Read in sample table
```{r,message=FALSE,warning=FALSE}
sample <- read.csv("totalRNA_sample_table.csv",header=T,row.names=1)
```

# Match reads count  and sample tables
```{r,message=FALSE,warning=FALSE,results='hide'}
sample <- sample[order(rownames(sample)),]
rna <- rna[,colnames(rna) %in% rownames(sample)]
rna <- rna[,order(colnames(rna))]
nrow(rna)
ncol(rna)
colnames(rna) == rownames(sample)
```

# Get vst data of all samples
```{r,message=FALSE,warning=FALSE}
dds <- DESeqDataSetFromMatrix(countData = rna,colData = sample,design = ~ Phase + Sex + probality + FinalDriverPathway.10Jan2020)
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq(dds,minReplicatesForReplace=Inf)
vst <- vst(dds,blind = F)
write.csv(assay(vst),"totalRNA_vst.csv")
```

# Read in tumor-normal paired sample table and match total RNA expression data
```{r,message=FALSE,warning=FALSE}
sample2 <- read.csv("paired_totalRNA_sample_table.csv",header=T,row.names=1)
sample2 <- sample2[order(rownames(sample2)),]
vst2 <- vst[,colnames(vst) %in% rownames(sample2)]
vst2
```
# Read in the k=2 clustering result of all tumor and normal samples using the vst expression data of top 1000 most variable genes with the tool the Consensus Cluster Plus
```{r,message=FALSE,warning=FALSE}
cl <- read.csv("totalRNA_tumor_normal_all_clust_CCP_top1000mostvariable_k2.csv",row.names=1)
cl2 <- cl[rownames(cl)%in%rownames(sample2),drop=F,]
cl2 <- cl2[order(rownames(cl2)),drop=F,]
vst2 $ cluster <- as.factor(cl2 $ x)
```

# PCA plot of the tumoar normal paired samples
```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(pheatmap)
library("vsn")
pdf("totalRNA_pca_plot_paired_samples.pdf")

pcaData <- plotPCA(vst2, intgroup=c("Specimen.Type.ID", "cluster"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))
ggplot(pcaData, aes(PC1, PC2, color=Specimen.Type.ID, shape=cluster)) +
  geom_point(size=2) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  ggtitle("tumor normal paired samples, total RNA")
dev.off()

```
