---
title: "PCA plots of tumor samples excluding samples with unknown drivers"
author: "Jieqiong Dai"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
      html_document:
        code_folding: hide
        toc: true
        toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Total RNA
## Reads in total RNA reads count data
```{r,message=FALSE,warning=FALSE}
library(DESeq2)
rna <- read.csv("totalRNA_reads_count_ready.csv",header=T,row.names=1)
```
## Read in sample table
```{r,message=FALSE,warning=FALSE}
sample <- read.csv("totalRNA_sample_table.csv",header=T,row.names=1)
```
## Match reads count and sample tables
```{r,message=FALSE,warning=FALSE,results='hide'}
sample <- sample[order(rownames(sample)),]
rna <- rna[,colnames(rna) %in% rownames(sample)]
rna <- rna[,order(colnames(rna))]
nrow(rna)
ncol(rna)
colnames(rna) == rownames(sample)
```
## Get vst data of all samples
```{r,message=FALSE,warning=FALSE}
dds <- DESeqDataSetFromMatrix(countData = rna,colData = sample,design = ~ Phase + Sex + probality + FinalDriverPathway.10Jan2020)
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq(dds,minReplicatesForReplace=Inf)
vst <- vst(dds,blind = F)
write.csv(assay(vst),"totalRNA_vst.csv")
```
## Read in tumor sample table (excluding samples with unkown drivers) and match total RNA expression data
```{r,message=FALSE,warning=FALSE,results='hide'}
sample2 <- read.csv("tumor_sample_drop_unkndown_drivers.csv",header=T,row.names=2)
sample2 <- sample2[order(rownames(sample2)),]
sample3 <- sample2[sample2$cluster_mRNA_TP_all_k5!=".",]
vst2 <- vst[,vst$TSS.Subject.ID %in% rownames(sample3) & vst$Specimen.Type.ID=="Primary Tumor"]
vst2
vst2 <- vst2[,order(vst2$TSS.Subject.ID)]
rownames(sample3)==vst2$TSS.Subject.ID
vst2 $ Cluster <- as.factor(sample3$cluster_mRNA_TP_all_k5)
vst2 $ Driver <- as.factor(sample3$finalDriverPathway_v35)
```
## PCA plot
```{r,message=FALSE,warning=FALSE}
library(ggplot2)
library(pheatmap)
library("vsn")
library(ggthemes)
library("scales")
stata_pal(scheme = "s2color")
show_col(stata_pal("s2color")(15))
pdf("totalRNA_pca_plot_tumors_exclude_unknown_driver.pdf",width = 10,height = 10)
pcaData <- plotPCA(vst2, intgroup=c("Driver", "Cluster"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

ggplot(pcaData, aes(PC1, PC2, color=Driver, shape=Cluster)) +
  geom_point(size=2) +
  scale_color_manual(values=c("#1a476f","#ffd200","#55752f","#e37e00","#bfa19c","#c10534","#938dd2","#cac27e")) +
  scale_shape_manual(values = c(1, 4,5,16,17)) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme(panel.background = element_rect(fill='white',color = 'black'),legend.title = element_text(size=10),legend.text = element_text(size=8),legend.key.size = unit(5,"mm")) +
  ggtitle("tumor samples exclude samples with unknown drivers, total RNA")

dev.off()
```

# Micro RNA

## Reads in miRNA reads count data
```{r,message=FALSE,warning=FALSE}
mi <- read.csv("miRNA_reads_count_ready.csv",header=T,row.names=1)
```
## Read in sample table
```{r,message=FALSE,warning=FALSE}
sample <- read.csv("miRNA_sample_table.csv",header=T,row.names=1)
```
## Match reads count and sample tables
```{r,message=FALSE,warning=FALSE,results='hide'}
sample <- sample[order(rownames(sample)),]
mi <- mi[,colnames(mi) %in% rownames(sample)]
mi <- mi[,order(colnames(mi))]
nrow(mi)
ncol(mi)
colnames(mi) == rownames(sample)
```
## Get get vst data for all samples
```{r,message=FALSE,warning=FALSE}
dds <- DESeqDataSetFromMatrix(countData = mi,colData = sample,design = ~ Phase + Sex + probality + FinalDriverPathway.10Jan2020)
keep <- rowSums(counts(dds)) >= 10
dds <- dds[keep,]
dds <- DESeq(dds,minReplicatesForReplace=Inf)
vst <- vst(dds,blind = F,nsub = 300)
write.csv(assay(vst),"miRNA_vst.csv")
```
## Get tumor sample table (exlucing samples with unknown drivers) and match miRNA expression data
```{r,message=FALSE,warning=FALSE,results='hide'}
sample3 <- sample2[sample2$cluster_miRNA_TP_all_k5!=".",]
vst2 <- vst[,vst$TSS.Subject.ID %in% rownames(sample3) & vst$Specimen.Type.ID=="Primary Tumor"]
vst2
vst2 <- vst2[,order(vst2$TSS.Subject.ID)]
rownames(sample3)==vst2$TSS.Subject.ID
vst2 $ Cluster <- as.factor(sample3$cluster_miRNA_TP_all_k5)
vst2 $ Driver <- as.factor(sample3$finalDriverPathway_v35)
```
## PCA plot
```{r,message=FALSE,warning=FALSE}
pdf("miRNA_pca_plot_tumors_exclude_unknown_driver.pdf",width = 10,height = 10)
pcaData <- plotPCA(vst2, intgroup=c("Driver", "Cluster"), returnData=TRUE)
percentVar <- round(100 * attr(pcaData, "percentVar"))

ggplot(pcaData, aes(PC1, PC2, color=Driver, shape=Cluster)) +
  geom_point(size=2) +
  scale_color_manual(values=c("#1a476f","#ffd200","#55752f","#e37e00","#bfa19c","#c10534","#938dd2","#cac27e")) +
  scale_shape_manual(values = c(1, 4,5,16,17)) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme(panel.background = element_rect(fill='white',color = 'black'),legend.title = element_text(size=10),legend.text = element_text(size=8),legend.key.size = unit(5,"mm")) +
  ggtitle("tumor samples exclude samples with unknown drivers, micro RNA")

dev.off()
```

# Methylation

## Reads in methylation data
```{r,message=FALSE,warning=FALSE}
library(data.table)
library(matrixStats)
system.time(beta_all <- (fread("normBeta.txt",header=T)))
beta_all <- as.data.frame(beta_all)
rownames(beta_all) <- beta_all $ TargetID
beta_all <- beta_all[,order(colnames(beta_all))]
```
## Read in tumor sample table (excluding samples with unknown drivers) and match methylation data
```{r,message=FALSE,warning=FALSE,results='hide'}
sample <- read.csv("pheno.csv",header=T,row.names=2)
sample3 <- sample2[sample2$cluster_meth_TP_all_k3!=".",]
sample3 <- sample3[rownames(sample3),]
sample <- sample[sample$patient_ID%in%rownames(sample3)&sample$sample=="T",]
sample <- sample[order(sample$patient_ID),]
sample$patient_ID==rownames(sample3)
sample3$id <- rownames(sample)
sample3 <- sample3[order(sample3$id),]
beta <- beta_all[,colnames(beta_all)%in% rownames(sample)]
beta <- beta[,order(colnames(beta))]
ncol(beta)
colnames(beta)==sample3$id
```
## PCA plot
```{r,message=FALSE,warning=FALSE}
library(ggfortify)
pca <- t(beta)
pca_data <- prcomp(pca)
data <- as.data.frame(pca_data $ x)
rownames(data)==sample3$id
data $ Cluster <- as.factor(sample3$cluster_meth_TP_all_k3)
data $ Driver <- as.factor(sample3$finalDriverPathway_v35)

pdf("methylation_pca_plot_tumors_exclude_unknown_driver.pdf",width = 10,height=10)
percentVar <- round(100 * pca_data$sdev^2 / sum( pca_data$sdev^2 ))

ggplot(data, aes(PC1, PC2, color=Driver, shape=Cluster)) +
  geom_point(size=2) +
  scale_color_manual(values=c("#1a476f","#ffd200","#55752f","#e37e00","#bfa19c","#c10534","#938dd2","#cac27e")) +
  scale_shape_manual(values = c(1,4,5)) +
  xlab(paste0("PC1: ",percentVar[1],"% variance")) +
  ylab(paste0("PC2: ",percentVar[2],"% variance")) + 
  coord_fixed() +
  theme(panel.background = element_rect(fill='white',color = 'black'),legend.title = element_text(size=10),legend.text = element_text(size=8),legend.key.size = unit(5,"mm")) +
  guides(color = guide_legend(order = 2),shape = guide_legend(order = 1)) +
  ggtitle("tumor samples exclude samples with unknown drivers, methylation")

dev.off()


```
